---
title: "Design your own race map"
author: "Reiko Okamoto"
format: html
editor: visual
---

Inspiration: <https://mrkyourmoment.com/en-ca/products/official-2025-tcs-sydney-marathon-print?variant=47047643070718>

## Step 1: Load necessary packages

```{r}
library(tidyverse)
library(here)
library(trackeR)
library(osmdata)
library(sf)
library(sysfonts)
library(showtext)
library(grid)
```

## Step 2: Wrangle running app data

Tips!

-   [Exporting GPX files from Strava](https://support.strava.com/hc/en-us/articles/216918437-Exporting-your-Data-and-Bulk-Export)
-   [Exporting GPX files from Runkeeper](https://help.runkeeper.com/en/hc/how-to-export-your-runkeeper-data)

### Load GPX file from Runkeeper

```{r}
# GPS coordinates captured at 4-second intervals
# points_df <- ...
```

### Convert coordinates to sf object

```{r}
# points_sf <- ...

# line_sf <- ...
```

## Step 3: Wrangle OSM data

### Define bounding box

```{r}
bbox <- c(
  xmin = -75.74498,
  ymin = 45.38766,
  xmax = -75.71863,
  ymax = 45.41050
)
```

### Build Overpass query

```{r}
query <- opq(bbox = bbox)
```

### Extract OSM data

```{r}
# from ottawa-map.qmd
# "major" roads
roads_a <- query |> 
  add_osm_feature(
    key = "highway",
    value = c(
      "motorway", "trunk", "primary",
      "motorway_link", "trunk_link", "primary_link"
    )
  ) |> 
  osmdata_sf()

# "minor" roads
roads_b <- query |> 
  add_osm_feature(
    key = "highway",
    value = c(
      "secondary", "tertiary", "unclassified", "residential",
      "secondary_link", "tertiary_link"
    )
  ) |> 
  osmdata_sf()
```

#### Crop sf objects

```{r}
bbox_sf <- st_as_sfc(st_bbox(bbox, crs = 4326))

roads_a_cropped <- st_crop(roads_a$osm_lines, bbox_sf)
roads_b_cropped <- st_crop(roads_b$osm_lines, bbox_sf)
```

**Posit Cloud users:** For some reason, cropping in Posit Cloud gives error: `Error in geos_op2_geom("intersection", x, y, ...) : st_crs(x) == st_crs(y) is not TRUE`. I'm not sure if this is the most graceful solution, but replace the previous code chunk with this one to bypass the error (we have to explicitly set the CRS of the OSM data).

```{r}
# bbox_sf <- st_as_sfc(st_bbox(bbox, crs = 4326))
# 
# roads_a_cropped <- roads_a$osm_lines |> st_set_crs(4326) |> st_crop(bbox_sf)
# roads_b_cropped <- roads_b$osm_lines |> st_set_crs(4326) |> st_crop(bbox_sf)
```

## Step 4: Visualize sf objects

```{r}
# define colour palette
roads_colour <- "#1c3c51"
route_colour <- "#ffffff"
bkgd_colour <- "#0D1E2A"
```

```{r}
# plt <- ...
```

## Step 7: Add text and export

```{r}
# define colour palette
# page_colour <- ...
# text_colour <- ...
# bib_no <- ...
# name <- ...
# distance <- ...
# time <- ...
# month_day <- ...
# year <- ...
```

```{r}
sysfonts::font_add_google(
  "Roboto Condensed",
  "Roboto Condensed",
  regular.wt = 400
)
png("race-map.png", width = 8, height = 10, units = "in", res = 300)
showtext::showtext_begin()

vp1 <- viewport(x = 0.5, y = 0.5, width = 1, height = 1)
pushViewport(vp1)
grid.rect(gp = gpar(fill = page_colour))
upViewport()

vp2 <- viewport(x = 0.5, y = 0.575, width = 0.9, height = 0.75)
pushViewport(vp2)
print(plt, newpage = FALSE)
upViewport()

vp3 <- viewport(x = 0.85, y = 0.15, width = 0.2, height = 0.05)
pushViewport(vp3)
grid.text(
  bib_no, gp = gpar(fontfamily = "Roboto Condensed", fontsize = 24)
)
upViewport()

vp4 <- viewport(x = 0.55, y = 0.15, width = 0.4, height = 0.05)
pushViewport(vp4)
grid.text(
  name, gp = gpar(fontfamily = "Roboto Condensed", fontsize = 30)
)
upViewport()

vp5 <- viewport(x = 0.65, y = 0.12, width = 0.6, height = 0.005)
pushViewport(vp5)
grid.rect(gp = gpar(fill = text_colour))
upViewport()

vp6 <- viewport(x = 0.85, y = 0.095, width = 0.2, height = 0.04)
pushViewport(vp6)
grid.text(
  distance, gp = gpar(fontfamily = "Roboto Condensed", fontsize = 24)
)
upViewport()

vp7 <- viewport(x = 0.65, y = 0.095, width = 0.2, height = 0.04)
pushViewport(vp7)
grid.text(
  time, gp = gpar(fontfamily = "Roboto Condensed", fontsize = 24)
)
upViewport()

vp8 <- viewport(x = 0.45, y = 0.095, width = 0.2, height = 0.04)
pushViewport(vp8)
grid.text(
  month_day, gp = gpar(fontfamily = "Roboto Condensed", fontsize = 24)
)
upViewport()

vp9 <- viewport(x = 0.85, y = 0.06, width = 0.2, height = 0.02)
pushViewport(vp9)
grid.text(
  "KM", gp = gpar(fontfamily = "Roboto Condensed", fontsize = 14)
)
upViewport()

vp10 <- viewport(x = 0.65, y = 0.06, width = 0.2, height = 0.02)
pushViewport(vp10)
grid.text(
  "TIME", gp = gpar(fontfamily = "Roboto Condensed", fontsize = 14)
)
upViewport()

vp11 <- viewport(x = 0.45, y = 0.06, width = 0.2, height = 0.02)
pushViewport(vp11)
grid.text(
  year, gp = gpar(fontfamily = "Roboto Condensed", fontsize = 14)
)
upViewport()

showtext::showtext_end()
dev.off()
```
